---
title: "Maps"
author: "Adam Kiehl"
date: "2022-10-03"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # ggplot and dplyr
library(maps) # map plotting data
library(mapproj) # coordinate projection
library(ggdendro) # simple map theme
library(viridis) # color palettes
library(plotly) # interactive plot effects
library(gridExtra) # grid plotting
```

`Data Source:` The data for the below plots were downloaded from the World University Ranking 2022-2023 open source data set on Kaggle.com (https://www.kaggle.com/datasets/4c5d5aecdebdec38eee4f9de4546e92868be0f4dc1b0351078b307acec179158?resource=download). World universities were scored and ranked based on their rankings in four categories: education quality, post-graduation employability, faculty quality, and research prestige. 
  
`Spatial Units:` Countries were the spatial units chosen for the plots below. University scores and category rankings were aggregated by country to match this unit. 
  
`Purpose:` The plots below were generated to convey a sense of holistic higher education quality in countries across the world. First and primarily, median university score per country was plotted to give the best overall sense of a country's collegiate quality. Next, each of the four category rankings used to compute this overall score (education quality, post-graduation employability, faculty quality, and research prestige) were aggregated by country and plotted as well. This serves to show how each country's specific collegiate strengths contribute to it's overall score and may illuminate some surprising aspects of the data.
  
`Decisions:` Several decisions were made in the process of the plotting this university ranking data. First, since the spatial composition of the map did not play a factor in the comparisons of interest to this problem, a simple, aesthetically pleasing projection similar to the Robinson projection was used, and a simple white theme was used to maximize the data$\colon$ink ratio. Next, median was chosen as a measure of central tendency over mean so that extremely prestigious outlying universities (such as Harvard) were not overly influential. Next, it was decided that each aggregated category ranking should be plotted in addition to the aggregated overall university score, since plotting different aspects of the data can tell different stories. For these plots, the legend scales and color palettes were reversed because, in these cases, a lower number is more desirable. It was additionally decided that the four category plots should be plotted in a 2-by-2 frame to emphasize their nature as categories contributing to a whole score. Finally, an interactive hover effect was added to the main plot to easily display the exact median university score for each country and to allow users to zoom in on smaller countries. 
  
```{r, warning=FALSE}
# read in university ratings data
univ_data <- read.csv('university.csv') %>%
  mutate(Education = as.numeric(Education.Rank),
         Employ = as.numeric(Employability.Rank),
         Faculty = as.numeric(Faculty.Rank),
         Research = as.numeric(Research.Rank)) %>%
  select(-c(Education.Rank,
            Employability.Rank,
            Faculty.Rank,
            Research.Rank))
# fix unaligned regions between data frames
univ_data$Location[which(univ_data$Location == 'Northern Cyprus')] <- 'Cyprus'
univ_data$Location[which(univ_data$Location == 'Slovak Republic')] <- 'Slovakia'
univ_data$Location[which(univ_data$Location == 'United Kingdom')] <- 'UK'

head(univ_data)

# read world map polygon data
world_map <- map_data('world') %>%
  # create variable to preserve order
  mutate(order = 1:n())
```

```{r}
# create aggregated scores data frame
univ_score <- univ_data %>%
  # aggregate scores by country
  group_by(Location) %>%
  # compute median scores
  summarize(Score = median(Score, na.rm = TRUE))

# left join world map data with university scores data
data1 <- world_map %>%
  merge(univ_score, 
        by.x = 'region', 
        by.y = 'Location', 
        all.x = TRUE) %>%
  # sort result by order variable
  arrange(order)

# build median scores choropleth plot
plt1 <- ggplot(data1,
              mapping = aes(x = long, 
                            y = lat,
                            group = group,
                            # add choropleth data
                            fill = Score,
                            # add text aesthetic for hover effect
                            text = paste(region, ': ', round(Score, 1), sep = ''))) +
  # draw map in white
  geom_polygon(color = 'white', size = .05) +
  # scale map to proper coordinates
  coord_map(xlim = c(-200, 200),
            ylim = c(-100, 100)) +
  # choose a minimal theme
  theme_dendro() +
  # choose color palette
  scale_fill_viridis(option = 'G') +
  # add legend title
  labs(fill = 'Median\nUniversity\nScore')

# add interactive hover effect
plt1 <- ggplotly(plt1, tooltip = 'text') %>%
  # add plot title and subtitle
  layout(title = list(text = paste0('<br>',
                                    'Median University Rating by Country',
                                    '<br>',
                                    '<sup>',
                                    'University rating based on education, employability, faculty, and research',
                                    '<sup>')))

# display first plot
plt1
```

```{r}
# create aggregated ranks data frame
univ_edu <- univ_data %>%
  # aggregate ranks by country
  group_by(Location) %>%
  # compute median education ranks
  summarize(Education = median(Education, na.rm = TRUE))

# left join world map data with university education ranks
data2 <- world_map %>%
  merge(univ_edu, 
        by.x = 'region', 
        by.y = 'Location', 
        all.x = TRUE) %>%
  # sort result by order variable
  arrange(order)

# build median education ranks choropleth plot
plt2 <- ggplot(data2,
              mapping = aes(x = long, 
                            y = lat,
                            group = group,
                            # add choropleth data
                            fill = Education)) +
  # draw map in white
  geom_polygon(color = 'white', size = .05) +
  # scale map to proper coordinates
  coord_map(xlim = c(-200, 200),
            ylim = c(-100, 100)) +
  # choose a minimal theme
  theme_dendro() +
  # choose color palette
  scale_fill_viridis(option = 'F', direction = -1) +
  # reverse order of legend scale
  guides(fill = guide_legend(reverse = FALSE)) +
  # add title and legend title
  labs(title = 'Education',
       fill = 'Median\nUniversity\nEducation\nRank')
```

```{r}
# create aggregated ranks data frame
univ_employ <- univ_data %>%
  # aggregate ranks by country
  group_by(Location) %>%
  # compute median employability ranks
  summarize(Employ = median(Employ, na.rm = TRUE))

# left join world map data with university employability ranks
data3 <- world_map %>%
  merge(univ_employ, 
        by.x = 'region', 
        by.y = 'Location', 
        all.x = TRUE) %>%
  # sort result by order variable
  arrange(order)

# build median employability ranks choropleth plot
plt3 <- ggplot(data3,
              mapping = aes(x = long, 
                            y = lat,
                            group = group,
                            # add choropleth data
                            fill = Employ)) +
  # draw map in white
  geom_polygon(color = 'white', size = .05) +
  # scale map to proper coordinates
  coord_map(xlim = c(-200, 200),
            ylim = c(-100, 100)) +
  # choose a minimal theme
  theme_dendro() +
  # choose color palette
  scale_fill_viridis(option = 'F', direction = -1) +
  # reverse order of legend scale
  guides(fill = guide_legend(reverse = FALSE)) +
  # add title and legend title
  labs(title = 'Employability',
       fill = 'Median\nUniversity\nEmployability\nRank')
```

```{r}
# create aggregated ranks data frame
univ_faculty <- univ_data %>%
  # aggregate ranks by country
  group_by(Location) %>%
  # compute median faculty ranks
  summarize(Faculty = median(Faculty, na.rm = TRUE))

# left join world map data with university faculty ranks
data4 <- world_map %>%
  merge(univ_faculty, 
        by.x = 'region', 
        by.y = 'Location', 
        all.x = TRUE) %>%
  # sort result by order variable
  arrange(order)

# build median faculty ranks choropleth plot
plt4 <- ggplot(data4,
              mapping = aes(x = long, 
                            y = lat,
                            group = group,
                            # add choropleth data
                            fill = Faculty)) +
  # draw map in white
  geom_polygon(color = 'white', size = .05) +
  # scale map to proper coordinates
  coord_map(xlim = c(-200, 200),
            ylim = c(-100, 100)) +
  # choose a minimal theme
  theme_dendro() +
  # choose color palette
  scale_fill_viridis(option = 'F', direction = -1) +
  # reverse order of legend scale
  guides(fill = guide_legend(reverse = FALSE)) +
  # add title and legend title
  labs(title = 'Faculty',
       fill = 'Median\nUniversity\nFaculty\nRank')
```

```{r}
# create aggregated ranks data frame
univ_research <- univ_data %>%
  # aggregate ranks by country
  group_by(Location) %>%
  # compute median faculty ranks
  summarize(Research = median(Research, na.rm = TRUE))

# left join world map data with university research ranks
data5 <- world_map %>%
  merge(univ_research, 
        by.x = 'region', 
        by.y = 'Location', 
        all.x = TRUE) %>%
  # sort result by order variable
  arrange(order)

# build median research ranks choropleth plot
plt5 <- ggplot(data5,
              mapping = aes(x = long, 
                            y = lat,
                            group = group,
                            # add choropleth data
                            fill = Research)) +
  # draw map in white
  geom_polygon(color = 'white', size = .05) +
  # scale map to proper coordinates
  coord_map(xlim = c(-200, 200),
            ylim = c(-100, 100)) +
  # choose a minimal theme
  theme_dendro() +
  # choose color palette
  scale_fill_viridis(option = 'F', direction = -1) +
  # reverse order of legend scale
  guides(fill = guide_legend(reverse = FALSE)) +
  # add title and legend title
  labs(title = 'Research',
       fill = 'Median\nUniversity\nResearch\nRank')
```

```{r}
# display category plots
grid.arrange(plt2, plt3, plt4, plt5,
             # define plotting grid
             ncol = 2,
             # add title to grid
             top = 'Median Category Rankings')
```
